### 2016-08-19 Supplement.tex
Added captions to all figures.  Rotated and split the large table of simulation parameter values.

\documentclass[useAMS,usenatbib,referee,12pt]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{parskip}
\usepackage[round]{natbib}
\usepackage{caption}
\usepackage{tabularx, booktabs}
\usepackage{float, placeins}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{rotating}

\newcommand{\pdet}{p^{(det)}}
\newcommand{\capt}[1]{Caterpillar plots of posterior parameter estimates from all models fit to the simulated #1 dataset (from simulations involving covariates and random effects).  Black and orange lines depict central 95\% and 50\% credible intervals, respectively.  The black dot is the posterior meadian.  The blue 'X' is the true parameter value and is positioned on the correctly specified model.  $\beta$'s are fixed effect parameters, $\sigma$'s are random effect standard deviations, and $\gamma$ is a mixing parameter.  Parameters designated with an 'A' are abundance parameters; those designated with a 'D' are detection parameters.}
\newcommand{\sfig}[2]{\begin{figure}
  \includegraphics[width=1.0\textwidth, page=#2]{Sims/SimFull/Posteriors_all_simdat.pdf}
  \caption{\capt{#1}}
  \end{figure}
}
\renewcommand{\thesection}{S-\arabic{section}}   % Label section with 'S-#'
\renewcommand{\thetable}{S-\arabic{table}}
\renewcommand{\thefigure}{S-\arabic{figure}}

\begin{document}

\section{Tables}
\begin{table}[!ht]
\centering
\begin{tabular}{lll|ccc}
  \hline
Family & Mixture & Peaked & $\gamma$ & $\varphi$ & $\alpha$ \\ 
  \hline
Exponential & Non-mixture & &  & -1.827 &  \\ 
  & Mixture & & 0.65 & -2.138 &  \\ 
  \hline
  Gamma & Non-mixture & Non-peaked &  & -3.279 & 0.257 \\ 
  & & Peaked &  & -0.746 & 3.371 \\ 
  & Mixture & Non-peaked & 0.65 & -2.773 & 0.577 \\ 
  & & Peaked & 0.65 & -1.210 & 2.491 \\ 
  \hline
  Lognormal & Non-mixture & Non-peaked &  & -0.341 & 2.330 \\ 
  & & Peaked &  & -1.872 & 0.512 \\ 
  & Mixture & Non-peaked & 0.65 & -1.462 & 1.674 \\ 
  & & Peaked & 0.65 & -1.992 & 0.618 \\ 
  \hline
  Weibull & Non-mixture & Non-peaked &  & -1.165 & 0.418 \\ 
  & & Peaked &  & -2.042 & 1.829 \\ 
  & Mixture & Non-peaked & 0.65 & -2.063 & 0.687 \\ 
  & & Peaked & 0.65 & -2.201 & 1.621 \\ 
   \hline
\end{tabular}
\caption{\label{simvals}Table of parameter values used to generate data for: (i) the mixture vs. non-mixture simulation, and (ii) the constant vs. non-constant rate simulation.  Here $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\varphi$ is the detection rate parameter, and $\alpha$ is the shape parameter.}
\end{table}

\begin{sidewaystable}[h!]\footnotesize
\centering
\begin{tabular}{lll|rrrr|rrrrr}
  \hline
Family & Mixture & Peak & $\beta^A_1$ & $\beta^A_2$ & $\beta^A_3$ & $\beta^A_4$ & $\beta^D_1$ & $\beta^D_2$ & $\beta^D_3$ & $\beta^D_4$ & $\beta^D_5$  \\
  \hline
Exponential & Non-mixture & &0.122&0.002&0.094&-1.039&-0.01&-0.056&0.118&0.249&0.164 \\
& Mixture & &0.124&0.012&0.064&-1.07&-0.068&-0.146&0.03&0.312&0.27 \\
Gamma & Non-mixture & Non-peaked &0.129&-0.001&0.098&-1.039&0.048&-0.104&-0.12&0.141&0.129 \\
& & Peaked &0.129&-0.001&0.098&-1.039&0.048&-0.104&-0.12&0.141&0.129 \\
& Mixture & Non-peaked &0.128&-0.005&0.079&-1.051&-0.048&-0.126&-0.042&0.21&0.2 \\
& & Peaked &0.128&-0.005&0.079&-1.051&-0.048&-0.126&-0.042&0.21&0.2 \\
Lognormal & Non-mixture & Non-peaked &0.126&0.031&0.096&-1.047&0.163&-0.099&-0.071&0.31&0.079 \\
& & Peaked &0.126&0.031&0.096&-1.047&0.163&-0.099&-0.071&0.31&0.079 \\
& Mixture & Non-peaked &0.126&0.017&0.07&-1.068&0.006&-0.14&-0.038&0.341&0.191 \\
& & Peaked &0.126&0.017&0.07&-1.068&0.006&-0.14&-0.038&0.341&0.191 \\
Weibull & Non-mixture & Non-peaked &0.129&0.011&0.102&-1.04&0.101&-0.104&-0.127&0.204&0.086 \\
& & Peaked &0.129&0.011&0.102&-1.04&0.101&-0.104&-0.127&0.204&0.086 \\
& Mixture & Non-peaked &0.129&-0.002&0.081&-1.049&-0.028&-0.122&-0.069&0.204&0.183 \\
& & Peaked &0.129&-0.002&0.081&-1.049&-0.028&-0.122&-0.069&0.204&0.183 \\
  \hline
\end{tabular}\\
\vspace{1.25cm}
\begin{tabular}{ll|lrrrrrrr}
  \hline
Family & Mixture & Peak & Intercept$^A$ & Intercept$^D$ & $\sigma_{A1}$ & $\sigma_{A2}$ & $\sigma_D$ & $\gamma$ & $\alpha$ \\
\hline
Exponential & Non-mixture & &0.975&-1.146&0.139&0.094&0.242& & \\
& Mixture & &1.097&-1.97&0.13&0.091&0.259&0.647& \\
Gamma & Non-mixture & Non-peaked &1.383&-3.941&0.121&0.072&0.39& & 0.316 \\
& & Peaked &1.134&-0.746&0.121&0.072&0.39& & 3.371 \\
& Mixture & Non-peaked &1.215&-2.7&0.121&0.079&0.332&0.79& 0.597 \\
& & Peaked &1.134&-1.21&0.121&0.079&0.332&0.65& 2.491 \\
Lognormal & Non-mixture & Non-peaked &1.613&-2.444&0.151&0.077&0.353& & 3.287 \\
& & Peaked &1.134&-1.872&0.151&0.077&0.353& & 0.512 \\
& Mixture & Non-peaked &1.282&-2.051&0.14&0.079&0.312&0.706& 1.467 \\
& & Peaked &1.134&-1.992&0.14&0.079&0.312&0.65& 0.618 \\
Weibull & Non-mixture & Non-peaked &1.57&-3.137&0.131&0.074&0.383& & 0.387 \\
& & Peaked &1.134&-2.042&0.131&0.074&0.383& & 1.829 \\
& Mixture & Non-peaked &1.294&-2.354&0.122&0.078&0.344&0.814& 0.653 \\
& & Peaked &1.134&-2.201&0.122&0.078&0.344&0.65& 1.621 \\
   \hline
\end{tabular}
\caption{\label{tbl:sim3}Table of parameter values used to generate data for simulations with covariates.  Here, $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\beta$'s are fixed effects, $\sigma$'s are random effect standard deviations, $\alpha$ is a shape parameter, and superscripts of $A$ and $D$ indicate abundance and detection parameters, respectively.}
\end{sidewaystable}
%\begin{sidewaystable}[ht]
%\centering
%\begin{tabular}{lllrrrrrrrrrrrrrrrr}
%  \hline
%Family & Mixture & Peak & Intercept$^A$ & $\beta^A_1$ & $\beta^A_2$ & $\beta^A_3$ & $\beta^A_4$ & Intercept$^D$ & $\beta^D_1$ & $\beta^D_2$ & $\beta^D_3$ & $\beta^D_4$ & $\beta^D_5$ & $\sigma_A[1]$ & $\sigma_A[2]$ & $\sigma_D$ & $\gamma$ & $\alpha$ \\ 
%  \hline
%Exponential & Non-mixture &  & 0.975 & 0.122 & 0.002 & 0.094 & -1.039 & -1.146 & -0.010 & -0.056 & 0.118 & 0.249 & 0.164 & 0.139 & 0.094 & 0.242 &  &  \\ 
%   & Mixture &  & 1.097 & 0.124 & 0.012 & 0.064 & -1.070 & -1.970 & -0.068 & -0.146 & 0.030 & 0.312 & 0.270 & 0.130 & 0.091 & 0.259 & 0.647 &  \\ 
%  Gamma & Non-mixture & Non-peaked & 1.383 & 0.129 & -0.001 & 0.098 & -1.039 & -3.941 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 0.316 \\ 
%   &  & Peaked & 1.134 & 0.129 & -0.001 & 0.098 & -1.039 & -0.746 & 0.048 & -0.104 & -0.120 & 0.141 & 0.129 & 0.121 & 0.072 & 0.390 &  & 3.371 \\ 
%   & Mixture & Non-peaked & 1.215 & 0.128 & -0.005 & 0.079 & -1.051 & -2.700 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.790 & 0.597 \\ 
%   &  & Peaked & 1.134 & 0.128 & -0.005 & 0.079 & -1.051 & -1.210 & -0.048 & -0.126 & -0.042 & 0.210 & 0.200 & 0.121 & 0.079 & 0.332 & 0.650 & 2.491 \\ 
%  Lognormal & Non-mixture & Non-peaked & 1.613 & 0.126 & 0.031 & 0.096 & -1.047 & -2.444 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 3.287 \\ 
%   &  & Peaked & 1.134 & 0.126 & 0.031 & 0.096 & -1.047 & -1.872 & 0.163 & -0.099 & -0.071 & 0.310 & 0.079 & 0.151 & 0.077 & 0.353 &  & 0.512 \\ 
%   & Mixture & Non-peaked & 1.282 & 0.126 & 0.017 & 0.070 & -1.068 & -2.051 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.706 & 1.467 \\ 
%   &  & Peaked & 1.134 & 0.126 & 0.017 & 0.070 & -1.068 & -1.992 & 0.006 & -0.140 & -0.038 & 0.341 & 0.191 & 0.140 & 0.079 & 0.312 & 0.650 & 0.618 \\ 
%  Weibull & Non-mixture & Non-peaked & 1.570 & 0.129 & 0.011 & 0.102 & -1.040 & -3.137 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 0.387 \\ 
%   &  & Peaked & 1.134 & 0.129 & 0.011 & 0.102 & -1.040 & -2.042 & 0.101 & -0.104 & -0.127 & 0.204 & 0.086 & 0.131 & 0.074 & 0.383 &  & 1.829 \\ 
%   & Mixture & Non-peaked & 1.294 & 0.129 & -0.002 & 0.081 & -1.049 & -2.354 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.814 & 0.653 \\ 
%   &  & Peaked & 1.134 & 0.129 & -0.002 & 0.081 & -1.049 & -2.201 & -0.028 & -0.122 & -0.069 & 0.204 & 0.183 & 0.122 & 0.078 & 0.344 & 0.650 & 1.621 \\ 
%   \hline
%\end{tabular}
%\begin{tabular}
%\end{tabular}
%\caption{\label{tbl:sim3}Table of parameter values used to generate data for simulations with covariates.  Here, $\gamma$ is a mixing parameter for the proportion of `hard to detect' individuals, $\beta$'s are fixed effects, $\sigma$'s are random effect standard deviations, $\alpha$ is a shape parameter, and superscripts of $A$ and $D$ indicate abundance and detection parameters, respectively.}
%\end{sidewaystable}
\FloatBarrier

\section{Figures}

\begin{figure}[h!]\centering
\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_correct.pdf}
\caption{\label{pdet_cater_correct} Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$.  
Inference models come from the same family as the dataset but may differ in the presence/absence of a mixture component.  
Each column presents one family of simulated dataset.  
Upper plots show non-mixture datasets; lower plots show mixture datasets.  
`X' marks the expected marginal probability of detection based on true parameter values.}
\includegraphics[width=0.98\textwidth]{Sims/SimFull/pdet_cater_family.pdf}
\caption{\label{pdet_cater_family}  Sim3 caterpillar plots of posterior 50\% and 95\% credible intervals for the marginal probability of detection $\pdet$. All data and inference models include mixtures.  
Each plot presents one simulated dataset.  
`X' marks the expected marginal probability of detection based on true parameter values.}
\end{figure}

\subsection{Posteriors parameter estimates from simulations including covariates and random effects}\label{s:sim3}
\sfig{non-mixture exponential}{1}
\sfig{exponential mixture}{2}
\sfig{nonpeaked non-mixture gamma}{3}
\sfig{nonpeaked gamma mixture}{4}
\sfig{peaked non-mixture gamma}{5}
\sfig{peaked gamma mixture}{6}
\sfig{nonpeaked non-mixture lognormal}{7}
\sfig{nonpeaked lognormal mixture}{8}
\sfig{peaked non-mixture lognormal}{9}
\sfig{peaked lognormal mixture}{10}
\sfig{nonpeaked non-mixture Weibull}{11}
\sfig{nonpeaked Weibull mixture}{12}
\sfig{peaked non-mixture Weibull}{13}
\sfig{peaked Weibull mixture}{14}

%\small
%\section{Stan Models}\label{s:models}
%\subsection{Exponential Model}
%\lstinputlisting[breaklines]{./OVEN/Expo_model.stan}
%\subsection{Exponential Mixture Model}
%\lstinputlisting[breaklines]{./OVEN/ExpoMix.stan}
%\subsection{Gamma Model}
%\lstinputlisting[breaklines]{./OVEN/Gamma_model.stan}
%\subsection{Gamma Mixture Model}
%\lstinputlisting[breaklines]{./OVEN/GammaMix.stan}
%\subsection{Lognormal Model}
%\lstinputlisting[breaklines]{./OVEN/LogNormal_model.stan}
%\subsection{Lognormal Mixture Model}
%\lstinputlisting[breaklines]{./OVEN/LogNormalMix.stan}
%\subsection{Weibull Model}
%\lstinputlisting[breaklines]{./OVEN/Weibull_model.stan}
%\subsection{Weibull Mixture Model}
%\lstinputlisting[breaklines]{./OVEN/WeibullMix.stan}

\end{document}